---
title: "KyleSha_L10_assignment"
author: "Kyle Sha"
date: "`r Sys.Date()`"
output:
  pdf_document:
    highlight: tango
    df_print: kable
  html_document:
    df_print: paged
geometry: margin=0.5in
fontsize: 10pt
usepackage: framed, xcolor, courier
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = TRUE, tidy = TRUE, fig.align = "center")

options(digits = 3, knitr.table.format = "latex")
setwd("~/")

library(tidyverse)
library(magrittr)
library(kableExtra)
```


```{r init, results=FALSE, fig.width=6, fig.height=6}
library("dbscan")
head(iris)
str(iris)
plot(iris)

# Do not use the outcome label (species name) in clustering
iris.m <- as.matrix(iris[,1:4])

# Make row.names pretty for easy debugging
row.names(iris.m) <- paste(row.names(iris),iris$Species)
speciesLabel <- stringr::str_extract(string=rownames(iris.m), pattern = "[a-z]{1,}")
```


## Q1. Perform k-means clustering on iris.m data set using 3 clusters
```{r Q1}
km <- kmeans(x = iris.m, centers = 3)
```


## Q2a. Using the results of k-means clustering: Determine a confusion matrix of species / cluster.
```{r Q2a}
confusion <- data.frame("species" = speciesLabel, "cluster"= km$cluster)

knitr::kable(x = table(confusion),
					 caption = "Confusion matrix for k-means.",
					 booktabs = TRUE, longtable = TRUE) %>%
kableExtra::kable_styling(latex_options = c("striped"), full_width = FALSE) %>%
kableExtra::row_spec(row = 0, bold = TRUE)
```


## Q2b. Assuming that a cluster 'belongs' to the most popular species in the cluster, what is the accuracy of the classification?

\vspace{3mm}\begin{leftbar}\textsf{\textbf{ANSWER Q2b}\\
The accuracy for k-means is (50+48+36)/150 = 89.3%
}\end{leftbar}\vspace{5mm}

## Q3. Perform hierarchical clustering on iris.m.
```{r Q3}
iris.m.scaled <- scale(x=iris.m, center = TRUE, scale=TRUE)
iris.m.scaled.dist <- dist(x = iris.m.scaled, method = "euclidean")
hc <- hclust(d = iris.m.scaled.dist, method = "complete")
plot(hc, labels = FALSE)
rect.hclust(tree=hc, k=3, border = c("red", "green", "blue"))
```

## Q4a. Using the results of hierarchical clustering: Choose a tree cut point, so that only 3clusters are generated.
```{r Q4a}
cut <- cutree(tree = hc, h=5)
```

## Q4b. Determine a confusion matrix of species / cluster.
```{r Q4b}
confusion_4b <- data.frame("species"=speciesLabel, "cluster"=cut)

knitr::kable(x = table(confusion_4b),
					 caption = "Confusion matrix for hierarchical clustering.",
					 booktabs = TRUE, longtable = TRUE) %>%
kableExtra::kable_styling(latex_options = c("striped"), full_width = FALSE) %>%
kableExtra::row_spec(row = 0, bold = TRUE)
```

## Q4c. What is the accuracy of the classification?
\vspace{3mm}\begin{leftbar}\textsf{\textbf{ANSWER Q4c}\\
The accuracy for hierarchical clustering is (49+21+48)/150 = 78.7%
}\end{leftbar}\vspace{5mm}

## Q5. Perform Density-based Clustering (DBSCAN) on iris.m, choosing parameters that result in 3 clusters (you might either seek 3 clusters or (for simplicity) assume that noise is one of the clusters).
```{r Q5}
db <- dbscan::dbscan(x=iris.m.scaled.dist, eps = 0.5, minPts = ncol(iris.m.scaled)+1, borderPoints = TRUE)
```


## Q6a.Using results of DBSCAN: Determine a confusion matrix of species/cluster.
```{r Q6a}
confusion_6a <- data.frame("species" = speciesLabel, "cluster"=db$cluster)

knitr::kable(x = table(confusion_6a),
					 caption = "Confusion matrix for DBSCAN.",
					 booktabs = TRUE, longtable = TRUE) %>%
kableExtra::kable_styling(latex_options = c("striped"), full_width = FALSE) %>%
kableExtra::row_spec(row = 0, bold = TRUE)
```


## Q6b. What is the accuracy of the classification?
\vspace{3mm}\begin{leftbar}\textsf{\textbf{ANSWER Q6b}\\
The accuracy for DBSCAN is (18+45+39)/150 = 68%
}\end{leftbar}\vspace{5mm}


## Q7. In the absence of the labels (but with knowledge that there are 3 different species), how would you decide which is the best algorithm?

\vspace{3mm}\begin{leftbar}\textsf{\textbf{ANSWER Q7}\\
I would use k-means clustering, since I know the number of groupings \emph{a priori}.
}\end{leftbar}\vspace{5mm}

